---
title: "Fandango-movie-ratings"
author: "Mai Le"
date: "6/11/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(dplyr) #grouping and manipulating data
library(ggplot2) #data visualizations
library(tidyr) #cleaning data
library(RColorBrewer) #color palettes
```

#1-Fandango Movie Ratings

```{r, include=FALSE,message=FALSE, warning=FALSE}
#Read data
library(fivethirtyeight)
data("fandango")
```

##1.1 Identify the Top 5 best rated and Top 5 worst rated movies in the dataset
To identify the "best" and "worst" rated movies, we calculated the average score for each movive, which is the mean normalized score across different rating websites.
```{r, message=FALSE, warning=FALSE}
#Create a new variable avg.rating which includes the average normalized score for each movie across different rating websites
ratings <- fandango %>% mutate(avg=(fandango_ratingvalue+rt_norm+rt_user_norm+metacritic_norm+metacritic_user_nom+imdb_norm)/6) 
```

Now we can sort the data by ascending or descending order to find the top 5 rated and bottom 5 rated movies. 
```{r, message=FALSE, warning=FALSE}
# Top 5 rated movies across all websites
ratings %>% arrange(desc(avg)) %>% select(1,24) %>% head(5)
```

```{r, message=FALSE, warning=FALSE}
# Bottom 5 rated movies across all websites
ratings %>% arrange(avg) %>% select(1,24) %>% head(5)
```




##1.2 Visualize the difference between Fandango stars and actual Fandango ratings
```{r, message=FALSE, warning=FALSE}
#transform data from wide to long format
ratings.wide <- fandango %>% gather(key=rating.type, value=score, fandango_stars,fandango_ratingvalue) 

# Density Area plot
ggplot(ratings.wide, aes(x=score, y=..density.., fill=rating.type)) +
  geom_area(stat ="bin",alpha=0.6) + theme_classic() 

ggplot(ratings.wide, aes(x= score, fill = rating.type)) + geom_density(alpha = 0.2)+ scale_fill_brewer(palette = "Set1")

ggplot(ratings.wide, aes(x= score, fill = rating.type)) + geom_density(stat= "bin",binwidth =0.5, alpha = 0.5) + labs(title ="Difference between Fandango stars and actual Fandango ratings", y = "Percentages", x="Star Ratings") + scale_fill_brewer(palette ="Set2")
```
The distribution of fandango_stars is more right skewed, indicating that there are more high ratings in fandango_stars than there are for fandango original ratings. 




##1.3 Top 5 movies that seem to appeal to critics but not the audience, and vice versa
We first compute a new variable CU.diff indicating the discrepancy between the average critic rating and the average user rating for each movie
```{r, message=FALSE, warning=FALSE}
# calculate average critic score for each film
ratings <- ratings %>% mutate(critic.avg = (rt_norm+metacritic_norm)/2)
# average audience score for each film
ratings <- ratings %>% mutate(user.avg = (fandango_ratingvalue+rt_user_norm+metacritic_user_nom+imdb_norm)/4)
# score difference between critic and user ratings for each film
ratings <- ratings %>% mutate(CU.diff = critic.avg-user.avg)
```

Sorting the data by CU. diff to find the top 5 movies that seem to appeal to critics but not the audience
```{r, message=FALSE, warning=FALSE}
# Top 5 movies that seem to appeal to critics but not the audience
ratings %>% arrange(desc(CU.diff)) %>% select(1,25,26,27) %>% head(5)
```
```{r, message=FALSE, warning=FALSE}
# Top 5 movies that users seem to like more than critics do
ratings %>% arrange(CU.diff) %>% select(1,25,26,27) %>% head(5)
```




##1.4 Alternative plots 
Taking inspirations from the fivethirtyeight article https://fivethirtyeight.com/features/fandango-movies-ratings/, let's make some alternative visualizations to the plots in the article. 

###1.4.1 Fandango's lopsided ratings curve
```{r, message=FALSE, warning=FALSE}
#subset data to include movies in theater in 2015 with 30+ reviews
fand2015 <- fandango[fandango$year == 2015 && fandango$fandango_votes+fandango$metacritic_user_vote_count >=30]

#convert fandango_stars to factor
fand2015$fand_stars_factor <-as.factor(fand2015$fandango_stars)

g1 = ggplot(fand2015) + 
  geom_bar(aes(x=fand_stars_factor, y=..count.., fill="light blue"), position = "dodge" ) + 
  theme_minimal() +
  theme(legend.position =  "none")

g1 + labs(title ="Fandango's Lopsided Ratings Curve", x = "Fandango stars", y = "Count") 
```

###1.4.2 Fandango Loves Movies
```{r, message=FALSE, warning=FALSE}
#transform data from wide to long format
ratings.w2<- fand2015 %>% gather(key=website, value=score, fandango_stars,rt_norm,rt_user_norm,metacritic_norm,metacritic_user_nom,imdb_norm) 

ratings.w2$website <- as.factor(ratings.w2$website)

# Side by side boxplots 
qplot(x=website, y=score, fill=website,geom='boxplot', data=ratings.w2)+ theme(legend.position="none")+ coord_flip() + scale_fill_brewer(palette ="Pastel1")
```

###1.4.3 Fandango Ratings Rounding
```{r, message=FALSE, warning=FALSE}
#transform rating_type to factor
ratings.wide$rating.type <- as.factor(ratings.wide$rating.type)

# Side by side boxplots 
qplot(x=rating.type, y=score, fill=rating.type,geom='boxplot', data=ratings.wide)+ theme(legend.position="none")+ coord_flip() + scale_fill_brewer(palette ="Pastel2")
```




##1.5 Average movie ratings over time for all the rating systems in the dataset
```{r, message=FALSE, warning=FALSE}
#transform data to long format 
ratings.15  <- fandango[fandango$year == 2015] %>% 
              mutate(rt.avg=mean(rt_norm))
ratings.14  <- fandango[fandango$year == 2014]


ratingwebs <- fandango %>% gather(key=website, value=score, rottentomatoes, rottentomatoes_user, metacritic)


qplot(x=year, y=avg, data=ratings)
```